|

### 2021-12-09 01-Chp-Background.Rmd
Range of small fixes (randomised, ApoE, etc)

A role for lipids in the aetiology of dementia is supported by both genetic linkage studies and functional cell biology studies. The generation of the amyloid plaques found in the brains of Alzheimer's patients is cholesterol dependent,,[@burns2003; @mizuno1999] while the most established genetic risk factor for late-onset dementia, apolipoprotein E (ApoE), is involved in cerebral cholesterol transport. Several other genes involved in cholesterol transport have also been found to be associated with increased AD susceptibility.[@beecham2014; @harold2009; @meng2007]

@@

Recent MR studies indicated that genetically determined low levels of LDL-c may cause a reduction in AD risk.[@larsson2017; @ostergaard2015] However, the effect was attenuated in sensitivity analysis that exclude the region surrounding the ApoE gene, the strongest known risk factor for Alzheimer's disease.[@kim2009] Inclusion of _ApoE4_ variants invalidates the exclusion restriction criteria (Assumption 3, above), as the risk reduction observed may be driven by variants in this region via a pathway independent of lipid levels. This was supported by further MR studies where _ApoE4_ variants were intentionally excluded.[@benn2017] Despite the increasing number of MR studies examining this topic, no systematic review of this study design as a source of evidence has been performed.

@@

I will attempt to obtain the raw data from relevant primary studies identified by the systematic review in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading). Any data obtained will be combined with that available from the DPUK portal as part of an IPD meta-analysis in Chapter \@ref(ipd-heading). This will enable the assessment of the effect of lipids on dementia stratified by key variables such as sex.

@@

To address this limitation, this thesis seeks to develop and apply a novel quantitative triangulation approach to answer causal questions on the effect of blood lipids on dementia outcomes. All existing evidence, regardless of study design, is first identified and assessed using domain-based tools by the systematic review presented in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading). The existing evidence base is then supplemented by two primary analyses, presented in Chapters \@ref(cprd-analysis-heading) and \@ref(ipd-heading). Finally all existing and new evidence is incorporated into the quantiatative triangulation framework in Chapter \@ref(tri-heading).

@@

-   **Chapter \@ref(background-heading):** Background information on dementia and blood lipid levels. This chapter provides an introduction to the topics covered in this thesis to non-subject area experts, and discusses the motivation for the remainder of the thesis.
-   **Chapter \@ref(sys-rev-tools-heading):** This chapter introduces a new tool, `medrxivr`, which was used to developed to allow for systematic searches of the health-related preprint repositories.
-   **Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading):** These chapters describe, respectively, the methods and results of a comprehensive systematic review and meta-analysis. This review examined all available evidence on the effect of blood lipids and interventions that modified blood lipids on dementia outcomes.
-   **Chapter \@ref(cprd-analysis-heading):** This chapter examines the relationship between lipid-regulating agent use and dementia outcomes in the Clinical Practice Research Datalink, a large primary care electronic health record database.
-   **Chapter \@ref(ipd-heading):** This chapter describes an individual participant data analysis of several longitudinal cohort studies to describe the relationship between blood serum lipids and dementia outcomes, stratified by important covariates such as sex.
-   **Chapter \@ref(tri-heading)**: This chapter integrates the evidence identified and produced by the preceeding chapters as part of a quantitative triangulation framework.
-   **Chapter \@ref(discussion-heading)**: This chapter summarise the key clinical and methodological findings of the thesis and discusses their implications for practice. The overall strengths and weaknesses of this project are discussed in detail, and further avenues of research are suggested.

@@

The experience of extracting data for the systematic review in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading) inspired a practical guide for researchers. This piece was co-written with Dr. Emily Hennessy (see Author Declarations in the front materials).

@@

The tool used to visualise the risk-of-bias assessments in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading) has been published in Research Synthesis Methods. See Appendix \@ref(appendix-robvis) for more details on this tool. Note that this publication does not describe the recently-developed functionality for producing bias direction plots, as described in Chapter \@ref(tri-heading).

@@

_"RoB 2.0: A revised tool to assess risk of bias in randomized trials"_: Webinar, co-presented with Dr. Theresa Moore as part of the Evidence Synthesis Ireland Methods Series.

>>>>

A role for lipids in the aetiology of dementia is supported by both genetic linkage studies and functional cell biology studies. The generation of the amyloid plaques found in the brains of Alzheimer's patients is cholesterol dependent,,[@burns2003; @mizuno1999] while the most established genetic risk factor for late-onset dementia, apolipoprotein E (_Apo_$\mathcal{E}$), is involved in cerebral cholesterol transport. Several other genes involved in cholesterol transport have also been found to be associated with increased AD susceptibility.[@beecham2014; @harold2009; @meng2007]

@@

Recent MR studies indicated that genetically determined low levels of LDL-c may cause a reduction in AD risk.[@larsson2017; @ostergaard2015] However, the effect was attenuated in sensitivity analysis that exclude the region surrounding the _Apo_$\mathcal{E}$ gene, the strongest known risk factor for Alzheimer's disease.[@kim2009] Inclusion of _Apo_$\mathcal{E}4$ variants invalidates the exclusion restriction criteria (Assumption 3, above), as the risk reduction observed may be driven by variants in this region via a pathway independent of lipid levels. This was supported by further MR studies where _Apo_$\mathcal{E}4$ variants were intentionally excluded.[@benn2017] Despite the increasing number of MR studies examining this topic, no systematic review of this study design as a source of evidence has been performed.

@@

I will attempt to obtain the raw data from relevant primary studies identified by the systematic review in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading). Any data obtained will be combined with that available from the DPUK portal as part of an IPD meta-analysis in Chapter \@ref(ipd-heading). This will enable the assessment of the effect of lipids on dementia stratified by key variables such as sex.

@@

To address this limitation, this thesis seeks to develop and apply a novel quantitative triangulation approach to answer causal questions on the effect of blood lipids on dementia outcomes. All existing evidence, regardless of study design, is first identified and assessed using domain-based tools by the systematic review presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading). The existing evidence base is then supplemented by two primary analyses, presented in Chapters \@ref(cprd-analysis-heading) and \@ref(ipd-heading). Finally all existing and new evidence is incorporated into the quantiatative triangulation framework in Chapter \@ref(tri-heading).

@@

-   **Chapter \@ref(background-heading):** Background information on dementia and blood lipid levels. This chapter provides an introduction to the topics covered in this thesis to non-subject area experts, and discusses the motivation for the remainder of the thesis.
-   **Chapter \@ref(sys-rev-tools-heading):** This chapter introduces a new tool, `medrxivr`, which was used to developed to allow for systematic searches of the health-related preprint repositories.
-   **Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading):** These chapters describe, respectively, the methods and results of a comprehensive systematic review and meta-analysis. This review examined all available evidence on the effect of blood lipids and interventions that modified blood lipids on dementia outcomes.
-   **Chapter \@ref(cprd-analysis-heading):** This chapter examines the relationship between lipid-regulating agent use and dementia outcomes in the Clinical Practice Research Datalink, a large primary care electronic health record database.
-   **Chapter \@ref(ipd-heading):** This chapter describes an individual participant data analysis of several longitudinal cohort studies to describe the relationship between blood serum lipids and dementia outcomes, stratified by important covariates such as sex.
-   **Chapter \@ref(tri-heading)**: This chapter integrates the evidence identified and produced by the preceeding chapters as part of a quantitative triangulation framework.
-   **Chapter \@ref(discussion-heading)**: This chapter summarise the key clinical and methodological findings of the thesis and discusses their implications for practice. The overall strengths and weaknesses of this project are discussed in detail, and further avenues of research are suggested.

@@

The experience of extracting data for the systematic review in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading) inspired a practical guide for researchers. This piece was co-written with Dr. Emily Hennessy (see Author Declarations in the front materials).

@@

The tool used to visualise the risk-of-bias assessments in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading) has been published in Research Synthesis Methods. See Appendix \@ref(appendix-robvis) for more details on this tool. Note that this publication does not describe the recently-developed functionality for producing bias direction plots, as described in Chapter \@ref(tri-heading).

@@

_"RoB 2.0: A revised tool to assess risk of bias in randomised trials"_: Webinar, co-presented with Dr. Theresa Moore as part of the Evidence Synthesis Ireland Methods Series.

### 2021-12-09 02-Chp-Systematic-Review-Tools.Rmd
Range of small fixes (randomised, ApoE, etc)

Searching these preprints as part of the systematic review described in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading) was a necessity, as many of the existing reviews on the topic of lipids and dementia have not considered this important source of evidence. At the time of writing, however, the bioRxiv and medRxiv websites allow only simple search queries as opposed to the complex searches containing Boolean logic operators (AND/OR/NOT) that information specialists use to query other major databases.[@bramer2018;@gusenbauer2020] Additionally, the best available extraction mechanism for obtaining references for all records returned by a search involved going through each record and downloading individual citations one-by-one. As the scale of these preprint databases increase, particularly in light of the massive expansion of the medRxiv repository as a result of COVID, this already time-consuming and error-prone method is no longer feasible.

@@

In addition to being used to search systematically health-related preprint servers, as illustrated in the systematic review presented in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading), `medrxivr` has other uses.

@@

-   I have outlined the motivation for developing this tool in relation to this thesis - more specifically, that it was used to perform systematic and reproducible searches of key literature sources used in the comprehensive systematic review described in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading).

>>>>

Searching these preprints as part of the systematic review described in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading) was a necessity, as many of the existing reviews on the topic of lipids and dementia have not considered this important source of evidence. At the time of writing, however, the bioRxiv and medRxiv websites allow only simple search queries as opposed to the complex searches containing Boolean logic operators (AND/OR/NOT) that information specialists use to query other major databases.[@bramer2018;@gusenbauer2020] Additionally, the best available extraction mechanism for obtaining references for all records returned by a search involved going through each record and downloading individual citations one-by-one. As the scale of these preprint databases increase, particularly in light of the massive expansion of the medRxiv repository as a result of COVID, this already time-consuming and error-prone method is no longer feasible.

@@

In addition to being used to search systematically health-related preprint servers, as illustrated in the systematic review presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading), `medrxivr` has other uses.

@@

-   I have outlined the motivation for developing this tool in relation to this thesis - more specifically, that it was used to perform systematic and reproducible searches of key literature sources used in the comprehensive systematic review described in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading).

### 2021-12-09 03-Chp-Systematic-Review-Methods.Rmd
Range of small fixes (randomised, ApoE, etc)

<!-- IDEA Stratify by _ApoE4_ genotype -->

@@

Eligible study designs were randomized controlled trials of LRA, cohort studies examining blood lipids or LRA, and genetic instrumental variable studies (more commonly known as Mendelian randomization studies) examining the effect of genetically increased/decreased blood lipid levels.

@@

I also searched clinical trial registries, for example ClinicalTrials.gov, to identify relevant randomized controlled trials. In addition, I searched the bioRxiv and medRxiv preprint repositories using the tool developed in Chapter \@ref(sys-rev-tools-heading) to identify potentially relevant preprinted studies (see Appendix \@ref(appendix-medrxivr-code) for the code used to search these preprint repositories).

@@

In line with best practice, where summary data were presented across two groups (e.g. age at baseline stratified by hypercholesterolemia status), the following approach was used to combine the groups:[@higgins2019]

@@

* bias arising from the randomization process, 
* bias due to deviations from intended intervention, 
* bias due to missing outcome data, 
* bias in measurement of the outcome, 
* bias in selection of the reported result.

@@

At present, no formalised risk-of-bias assessment tool for Mendelian randomization studies is available. Assessment of the risk of bias in Mendelian randomization studies was informed by the approach used in a previous systematic review of Mendelian randomisation,[@mamluk2020] as identified by a review of risk-of-bias assessments in systematic reviews of Mendelian randomisation studies.[@spiga2021] Advanced results from this review were obtained from contact with the review authors. A copy of this tool is available in Appendix \@ref(appendix-mr-rob), but in summary, results were assessed for bias arising from weak instruments, genetic and other confounding, pleiotropy and population stratification. Acceptable judgements for each of the five domains in the tool were "low", "moderate" and "high" risk of bias.

>>>>

<!-- IDEA Stratify by _Apo_$\mathcal{E}4$ genotype -->

@@

Eligible study designs were randomised controlled trials of LRA, cohort studies examining blood lipids or LRA, and genetic instrumental variable studies (more commonly known as Mendelian randomisation studies) examining the effect of genetically increased/decreased blood lipid levels.

@@

I also searched clinical trial registries, for example ClinicalTrials.gov, to identify relevant randomised controlled trials. In addition, I searched the bioRxiv and medRxiv preprint repositories using the tool developed in Chapter \@ref(sys-rev-tools-heading) to identify potentially relevant preprinted studies (see Appendix \@ref(appendix-medrxivr-code) for the code used to search these preprint repositories).

@@

In line with best practice, where summary data was presented across two groups (e.g. age at baseline stratified by hypercholesterolemia status), the following approach was used to combine the groups:[@higgins2019]

@@

* bias arising from the randomisation process, 
* bias due to deviations from intended intervention, 
* bias due to missing outcome data, 
* bias in measurement of the outcome, 
* bias in selection of the reported result.

@@

At present, no formalised risk-of-bias assessment tool for Mendelian randomisation studies is available. Assessment of the risk of bias in Mendelian randomisation studies was informed by the approach used in a previous systematic review of Mendelian randomisation,[@mamluk2020] as identified by a review of risk-of-bias assessments in systematic reviews of Mendelian randomisation studies.[@spiga2021] Advanced results from this review were obtained from contact with the review authors. A copy of this tool is available in Appendix \@ref(appendix-mr-rob), but in summary, results were assessed for bias arising from weak instruments, genetic and other confounding, pleiotropy and population stratification. Acceptable judgements for each of the five domains in the tool were "low", "moderate" and "high" risk of bias.
